More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2018-02-12 2016-Weak-FH-Bisim/WeakBisim.tex
added a todo
removed magical op new def of leaves new tr2 took guard into account until page 20

\LUDO{introduce magic op}
$\CreateISet{a}{i}$ creates a single entry indexed set $(i\mapsto a)$ \emph{but only if 
$a$ is not already an indexed set.}

@@

\item
The set of holes of a pNet is defined inductively; the sets of holes
in a pNet node and its subnets are all disjoint:
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset\\
\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
=J\uplus{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  \Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined 
inductively as:
\TODO{Nicer here at least}
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\!\mylangle S,s_0, \to\myrangle\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\overline{\Sort}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
I}\CreateISet{\Leaves(\pNet_i)}{i} }
\end{array}\]
\end{itemize}
\end{definition}

@@

@@

  For any pNet p, let $\Leaves(p) = \mylangle S_i,{s_i}_0, \to_i\myrangle^{i \in L}$ be 
  the set of pLTS at its leaves,
  then $States(p) = \{\triangleleft s_i^{i\in L}
  \triangleright| \forall i\in L. s_i \in S_i\}$.
\TODO{this definition is not valid because L is undefined and leaves only exist for pNets 
-> REMOVE? this would mean that a pNet is not a pLTS ...}
A pLTS being its own single leave:
  $States(\mylangle S,s_0, \to\myrangle) = \{\triangleleft s \triangleright| s \in S\}$.

@@

\paragraph{Predicates:}
Let
$\mylangle\overline{\pNet},\overline{\Sort},\symb{SV}_k^{k\in K} \myrangle$
be a pNet. Consider a synchronisation vector $SV_k$, for $k\in K$. We define a
predicate $\Pred$ relating
the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
\[\Pred_{\symb{sv}}(SV_k, a_i^{i\in I}, b_j^{j\in J}, v)\Leftrightarrow \bigg(
\begin{array}{l}
\exists {(a'_i)}^{i\in I},
{(b'_j)}^{j\in J},v'.\, SV_k={(a'_i)}^{i\in I}, {(b'_j)}^{j\in J}\rightarrow v'
\\~~\land
\forall i\in I.\, a_i=a'_i\land \forall j \in J.\, b_j=b'_j \land v=v'
\end{array} \bigg)\]

@@

	The second rule deals with pNet nodes: for each possible
	synchronisation vector applicable to the rule subject, the premisses
	include one {\em open transition} for each sub-pNet involved, one possible
	{\em action} for each Hole involved, and the predicate relating these
	with the resulting action of the vector.
	A key to understand this rule is that the open transitions are
	expressed in terms of the leaves and holes of the pNet structure,
	i.e. a flatten view of the pNet: e.g. $L$ is the index set of the
	Leaves, $L_m$ the index set of the leaves of one subnet indexed $m$, so all $L_m$
	are disjoint subsets of $L$. Thus the states in the open transitions,
	at each level, are tuples including states of all the
	leaves of the pNet, not only those involved in the chosen
	synchronisation vector.
	\LUDO{now with the magical operator ... $s_m$ and $s_m'$ are a bit ill-typed terms. 
	Not very happy of the new rule}
	\begin{description}
		\item[{\bf Tr2:}]
	\end{description}

	\noindent
    $\inferrule
    {k\!\in\! K \\ SV\!\!= \!{\tt clone}(SV_k) \!=\! \alpha_m^{m \in I_k\uplus J_k} \!\to\! 
    \alpha'_k \\
    	Leaves(p) \!=\! \pLTS_l^{l\in L} \\    	
    	\forall m\in I_k.
    	\pNet_m \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}}
    	%\land
    	%Leaves(\pNet_m) = \overline{\pLTS}^{L_k})
    	\\\forall m\in I_k. \CreateISet{s_m}{i}=s_{i}^{i \in L_m} \\
		\forall m\in I_k. \CreateISet{s_m'}{i}=(s_i^\prime)^{i\in L_m} \\
    	I' = \biguplus_{m\in I_k}\!\! I_m'
    	\\ J' = \biguplus_{m\in I_k}\!\! J'_m \uplus J_k 	\\
    	\Pred = \bigwedge_{m\in I_k}\!\! \Pred_m \land
    	\Pred_{\symb{sv}}(SV,a_i^{i\in I_k},b_j^{j\in J_k},v)\\
    	\forall j\!\!\in\!\! J_k. {\tt 
    		fresh}(b_j) \\ {\tt fresh}(v) \\ 
    	\forall i\in
    	L\backslash I'.\,s'_i=s_i 
    }
    {\mylangle \pNet_i^{i\in I}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\inferrule*[myfraction=\reddottedrule]
    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {v}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
    $

	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}

@@

\begin{lemma}[OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose:
	\[ P[Q]_{j_0}  
		\models
		{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I},
				\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {v}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $I\cap L_Q\neq 
		\emptyset$, i.e. $Q$ takes part in the reduction.
		 Then there exist $v_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {v}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{\,i\in I\cap L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {v_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred\subst{v_Q\leftarrow b_{j_0}} =(\Pred\,'
		\land \Pred\,'')$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

   \begin{small}\begin{proof}
       Consider each premise of the open transition (as constructed by rule TR2 in Definition~\ref{def:operationalSemantics}). 
We know each premise is true for $P[Q]_{j_0}$ and try to prove the equivalent premise for 
$P$. 
First, $K$ and the synchronisation vector $SV_k$ are unchanged\footnotemark (however 
$j_0$ passes from 
the set of subnets to the set of holes). Then $SV=clone(\alpha_j^{j\in I_k\uplus\{j_0\}\uplus 
J_k})$. $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$. 
\TODO{Revoir la def des leaves set ou multisets}

Now focus on OTs of the subnets\footnotemark[\thefootnote]:\\[-2ex]
\[\forall m\in I_k\cup \{j_0\}.
\pNet_m \models
\inferrule*[myfraction=\reddottedrule]
{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m},
	\{\xrightarrow{b_{j}}_j\}^{j\in J_m}, \Pred_m, \Post_m}
{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
	\ostate{s_{i}'^{\, i \in L_m}}}\]
Only elements of $I_k$ are useful to assert the premise for reduction of $P$; the last 
one ensures (note that $Q$ is at place $j_0$, and by definition of the open transition 
for $P[Q]_{j_0}$, $I_{j_0}=I\cap L_Q$, 
$L_{j_0}=L_Q$):\\[-2ex]
	\[Q\models{\inferrule*[myfraction=\reddottedrule]
		{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L_Q},
			\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred_{j_0},  
			\Post\,''}
		{\ostate{s_i^{i\in L_Q}} \OTarrow {v_{j_0}}
			\ostate{s_i'^{\, i\in L_Q}}}
	}\]
\TODO{We need to add: Note that $J_{j_0}=	J\cap\Holes(Q)$ this is a direct consequence of Tr2??}

This already ensures the second part of the conclusion if we 
choose\footnotemark[\thefootnote]  $v_Q=v_{j_0}$ ($\Pred\,''= \Pred_{j_0}$). Now let $I'= 
\biguplus I'_m=I\setminus L_Q$, $J'=\biguplus J'_m\uplus J_k 
\uplus\{j_0\}=J\setminus\Holes(Q)\uplus \{j_0\}$; the predicate is 
$\Pred\,'=\bigwedge_{m\in I_k}\Pred_m 
\land \Pred(SV,a_i^{i\in I_k},b_j^{j\in J_k\cup \{j_0\}},v)$
where\footnotemark[\thefootnote] $\Pred(SV,a_i^{i\in I_k},b_j^{j\in 
J_k},v)\Leftrightarrow 
\forall i\in I_k.\, \alpha_i=a_i\land \forall j \in J_k\cup \{j_0\}.\, \alpha_j=b_j \land 
v=\alpha'_k 
$. Modulo renaming of fresh variables, this is identical to the predicate that occurs in 
the source open transition except $\alpha_{j_0}=v_{j_0}$ has been replaced by  
$\alpha_{j_0}=b_{j_0}$.
Thus, $\Pred\subst{v_Q\leftarrow b_{j_0}}=(\Pred\,' 
\land \Pred\,'')$. %\TODO{the argument on renaming has mostly been
                 %removed is it still OK? yes.}. 
Finally, Post 
%conditions are easily split\footnote{\TODO{moved. could
%    even remove...?}Note that if the rules are not simplified, no
%  post-condition range over the variables of $P$ and $Q$ at the same
%  time} 
into conditions of the context $P$ and the pNet $Q$ (they are
builts similarly as they only deal with  
leaves): $\Post=\Post\,'\uplus \Post\,''$. This concludes the 
proof as we checked all the premises of the open transition for both $P$ and $Q$.\qed
\footnotetext{Cloning and freshness introduce alpha-conversion at many points 
of the proof; we 
	only 
	give major arguments concerning alpha-conversion to make the proof readable; in 
	general, fresh variables appear in each transition inside
        terms $b_j$, $v$, and 
	$\Pred$.}
\end{proof}
   \end{small}

>>>>

%\LUDO{introduce magic op}
%$\CreateISet{a}{i}$ creates a single entry indexed set $(i\mapsto a)$ \emph{but only if 
%$a$ is not already an indexed set.}

@@

\item
The set of holes of a pNet is defined inductively; the sets of holes
in a pNet node and its subnets are all disjoint:
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset\\
\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
=J\uplus{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  \Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined 
inductively as:
\TODO{Nicer here at least}
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\!\{\mylangle S,s_0, \to\myrangle\}\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\overline{\Sort}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
I}\Leaves(\pNet_i)\uplus\{i\mapsto \pNet_i|\pNet_i=\pLTS\}}
\end{array}\]
\end{itemize}
\end{definition}

@@

\TODO{update pred with guards}

@@

  For any pNet p, let $\Leaves(p) = \mylangle S_i,{s_i}_0, \to_i\myrangle^{i \in L}$ be 
  the set of pLTS at its leaves,
  then $States(p) = \{\triangleleft s_i^{i\in L}
  \triangleright| \forall i\in L. s_i \in S_i\}$.
A pLTS being its own single leave:
  $States(\mylangle S,s_0, \to\myrangle) = \{\triangleleft s \triangleright| s \in S\}$.

@@

\paragraph{Predicates:}
Let
$\mylangle\overline{\pNet},\overline{\Sort},\overline{\symb{SV}}\myrangle$
be a pNet. Consider a synchronisation vector $\symb{SV}\in \overline{\symb{SV}}$. We 
define a
predicate $\Pred$ relating
the actions of the involved sub-pNets and the resulting actions. This predicate verifies:
\[\Pred_{\symb{sv}}(SV, a_i^{i\in I}, b_j^{j\in J}, v)\Leftrightarrow \bigg(
\begin{array}{l}
\exists {(a'_i)}^{i\in I},
{(b'_j)}^{j\in J},v'.\, SV={(a'_i)}^{i\in I}, {(b'_j)}^{j\in J}\rightarrow v' | g
\\~~\land
\forall i\in I.\, a_i=a'_i\land \forall j \in J.\, b_j=b'_j \land v=v' \land g
\end{array} \bigg)\]
\LUDO{Is and g sufficient? I would say yes}

@@

	The second rule deals with pNet nodes: for each possible
	synchronisation vector applicable to the rule subject, the premisses
	include one {\em open transition} for each sub-pNet involved, one possible
	{\em action} for each Hole involved, and the predicate relating these
	with the resulting action of the vector.
	A key to understand this rule is that the open transitions are
	expressed in terms of the leaves and holes of the pNet structure,
	i.e. a flatten view of the pNet: e.g. $L$ is the index set of the
	Leaves, $L_m$ the index set of the leaves of one subnet indexed $m$, so all $L_m$
	are disjoint subsets of $L$. Thus the states in the open transitions,
	at each level, are tuples including states of all the
	leaves of the pNet, not only those involved in the chosen
	synchronisation vector.
	\LUDO{now with two cases of induction} \LUDO{Why I'm and not Im?}
	\begin{description}
		\item[{\bf Tr2:}]
	\end{description}

	\noindent
    $\inferrule
    {k\!\in\! K \\SV_k \!=\! \alpha_m^{m \in I\uplus J} \!\to\! 
    \alpha' |g \\
    	Leaves(\mylangle \pNet_i^{i\in I}, \overline{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \\    	
    	\forall m\!\!\in\!\! I. 	
    {\left(\begin{array}{l}
	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
		{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I'_m=\{m\} \land J'_m=\emptyset}
\end{array}\right)}\\
    	%\land
    	%Leaves(\pNet_m) = \overline{\pLTS}^{L_k})  	
     J' = \biguplus_{m\in I}\!\! J'_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I}\!\! \Pred_m \land
    	\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J},v)\\ 
    		I' = \biguplus_{m\in I}\!\! I_m'
    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
    {\tt fresh}(\alpha_m,\alpha',b_j,v) 
    }
    {\mylangle \pNet_i^{i\in I}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\inferrule*[myfraction=\reddottedrule]
    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {v}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
    $
	\medskip
%        \TODO{may be explain how $\Pred(SV,a_i^{i\in I_k},b_j^{j\in
%            J_k},v)$ is built ? You mean more than what is written on previous page????}
	%%    \TODO{I have tentatively added the sort constraint on hole actions, that was
	%%not included in the first version... I'm unsure whether this is the best place to
	%%include it, because it may change the decidability conditions on predicates}

@@

\begin{lemma}[OT decomposition]\label{lem-decompose} 
	Let $\Leaves(Q)=p_l^{l\in L_Q}$; suppose\footnote{Note that the hypotheses of the 
	lemma imply that $Q$ is 
	not a \pLTS\ but a similar lemma can be proven for a pLTS $Q$}:
	\[ P[Q]_{j_0}  
		\models
		{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I},
				\{\xrightarrow{b_j}_j\}^{j\in J}, \Pred,  
				\Post}
			{\ostate{s_i^{i\in L}} \OTarrow {v}
				\ostate{s_i'^{\, i\in L}}}
		}
	\]
		with  $J\cap\Holes(Q)\neq\emptyset$ or $I\cap L_Q\neq 
		\emptyset$, i.e. $Q$ takes part in the reduction.
		 Then there exist $v_Q$, $\Pred\,'$, $\Pred\,''$, 
		$\Post\,'$, $\Post\,''$ s.t.:\\[-2ex]
		%\[
		\begin{mathpar}
		P\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\setminus L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in (J\setminus \Holes(Q)) \cup \{j_0\}}, 
				\Pred\,',  
				\Post\,'}
			{\ostate{s_i^{i\in L\setminus L_Q}} \OTarrow {v}
				\ostate{s_i'^{\,i\in L\setminus L_Q}}}
		}%\]
	\vspace{-2.2ex}\\\text{and~~}
		%\[
		Q\models{\inferrule*[myfraction=\reddottedrule]
			{\{s_i\xrightarrow{a_i}_i s_i'\}^{\,i\in I\cap L_Q},
				\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred\,'',  
				\Post\,''}
			{\ostate{s_i^{i\in L_Q}} \OTarrow {v_Q}
				\ostate{s_i'^{\,i\in L_Q}}}
		}%\]
		\end{mathpar}
		and  $\Pred\subst{v_Q\leftarrow b_{j_0}} =(\Pred\,'
		\land \Pred\,'')$, $\Post=\Post\,'\uplus 
		\Post\,''$ where $\Post\,''$ is the restriction of $\Post$ over variables of 
		$\Leaves(Q)$.
\end{lemma}

   \begin{small}\begin{proof}
       Consider each premise of the open transition (as constructed by rule TR2 in 
       Definition~\ref{def:operationalSemantics}). 
	\begin{description}
		\item[{\bf Tr2:}]
	\end{description}

	\noindent
    $\inferrule
    {k\!\in\! K \\SV_k \!=\! \alpha_m^{m \in I\uplus J} \!\to\! 
    \alpha' |g \\
    	Leaves(\mylangle \pNet_i^{i\in I}, \overline{\Sort}, \symb{SV}_k^{k\in 
    	K}\myrangle) \!=\! \pLTS_l^{l\in L} \\    	
    	\forall m\!\!\in\!\! I. 	
    {\left(\begin{array}{l}
	{\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\lor }\\
		{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I'_m=\{m\} \land J'_m=\emptyset}
\end{array}\right)}\\
    	%\land
    	%Leaves(\pNet_m) = \overline{\pLTS}^{L_k})  	
     J' = \biguplus_{m\in I}\!\! J'_m \uplus J 	\\
    	\Pred = \bigwedge_{m\in I}\!\! \Pred_m \land
    	\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J},v)\\ 
    		I' = \biguplus_{m\in I}\!\! I_m'
    	\\\forall i\in	L\backslash I'.\,s'_i=s_i \\
    {\tt fresh}(\alpha_m,\alpha',b_j,v) 
    }
    {\mylangle \pNet_i^{i\in I}, \overline{\Sort}, \symb{SV}_k^{k\in K}\myrangle
    	\models
    	{\inferrule*[myfraction=\reddottedrule]
    		{\{s_i\xrightarrow{a_i}_i s_i^{\prime}\}^{i\in I^\prime},
    		\{\xrightarrow{b_j}_j\}^{j\in J^\prime}, \Pred, \uplus_{m\in I_k} 
    		\Post_m}
    		{\ostate{s_i^{i\in L}} \OTarrow {v}
    			\ostate{(s_i^\prime)^{i\in L}}}
    	}
    }
    $
	\medskip

We know each premise is true for $P[Q]_{j_0}$ and try to prove the equivalent premise for 
$P$. 
First, $K$ and the synchronisation vector $SV_k$ are unchanged\footnotemark (however 
$j_0$ passes from 
the set of subnets to the set of holes). 
%Then $SV=clone(\alpha_j^{j\in I_k\uplus\{j_0\}\uplus J_k})$.
We have $\Leaves(P[Q]_{j_0})=\Leaves(P)\uplus \Leaves(Q)$.

Now focus on OTs of the subnets. For each $m\in I\cup \{j_0\}$ we have one of the two 
following OT\footnotemark[\thefootnote]:\\[-2ex]
\[
\pNet_m \models\inferrule*[myfraction=\reddottedrule]
    	{\{s_{i}\xrightarrow{a_{i}}_i s_{i}'\}^{i\in I_m^\prime},
    	\{\xrightarrow{b_{j}}_j\}^{j\in J'_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {v_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\]
Or
\[{ \pNet_m 
    	 \models
    	\inferrule*[myfraction=\reddottedrule]
    	{\{s_m \xrightarrow{v_m} s_m'\},\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {v_m}
    		\ostate{s_m'}} \land I'_m=\{m\} \land J'_m=\emptyset}\]

Only elements of $I$ are useful to assert the premise for reduction of $P$; the last 
one ensures (note that $Q$ is at place $j_0$, and by definition of the open transition 
for $P[Q]_{j_0}$, $I'_{j_0}=I\cap L_Q$, 
$L_{j_0}=L_Q$, and $J_{j_0}=	J\cap\Holes(Q)$):\\[-2ex]
	\[Q\models{\inferrule*[myfraction=\reddottedrule]
		{\{s_i\xrightarrow{a_i}_i s_i'\}^{i\in I\cap L_Q},
			\{\xrightarrow{b_j}_j\}^{j\in J\cap\Holes(Q)}, \Pred_{j_0},  
			\Post\,''}
		{\ostate{s_i^{i\in L_Q}} \OTarrow {v_{j_0}}
			\ostate{(s_i')^{\, i\in L_Q}}}
	}\]

This already ensures the second part of the conclusion of the lemma, i.e. the OT for $Q$ 
if we 
choose\footnotemark[\thefootnote]  $v_Q=v_{j_0}$ ($\Pred\,''= \Pred_{j_0}$). Considering 
the OT of $P$ let $I'= 
\biguplus I'_m=I\setminus L_Q$, $J'=\biguplus J'_m\uplus J
\uplus\{j_0\}=J\setminus\Holes(Q)\uplus \{j_0\}$;  the predicate is 
$\Pred\,'=\bigwedge_{m\in I_k}\Pred_m  \land \Pred_{\symb{sv}}(SV_k,v_m^{m\in 
I},b_j^{j\in J\cup \{j_0\}},v)$
where\footnotemark[\thefootnote] $\Pred_{\symb{sv}}(SV_k,v_m^{m\in I},b_j^{j\in J\cup 
\{j_0\}},v)\Leftrightarrow 
\forall i\in I_k.\, \alpha_i=a_i\land \forall j \in J\cup \{j_0\}.\, \alpha_j=b_j \land 
v=\alpha'
$. Modulo renaming of fresh variables, this is identical to the predicate that occurs in 
the source open transition except $\alpha_{j_0}=v_{j_0}$ has been replaced by  
$\alpha_{j_0}=b_{j_0}$.
Thus, $\Pred\subst{v_Q\leftarrow b_{j_0}}=(\Pred\,' 
\land \Pred\,'')$. %\TODO{the argument on renaming has mostly been
                 %removed is it still OK? yes.}. 
Finally, Post 
%conditions are easily split\footnote{\TODO{moved. could
%    even remove...?}Note that if the rules are not simplified, no
%  post-condition range over the variables of $P$ and $Q$ at the same
%  time} 
into conditions of the context $P$ and the pNet $Q$ (they are
builts similarly as they only deal with  
leaves): $\Post=\Post\,'\uplus \Post\,''$. This concludes the 
proof as we checked all the premises of the open transition for both $P$ and $Q$.\qed
\footnotetext{Cloning and freshness introduce alpha-conversion at many points 
of the proof; we 
	only 
	give major arguments concerning alpha-conversion to make the proof readable; in 
	general, fresh variables appear in each transition inside
        terms $b_j$, $v$, and 
	$\Pred$.}
\end{proof}
   \end{small}
